\racets: Faceted Execution in Racket

### 2018-07-24 paper/paper.tex
taking out potentialy
adding cites to hygienic macros

\emph{Faceted execution} (FE) is a highly expressive (but potentially expensive)
language semantics enabling policy-agnostic programming. In FE, dynamic information-flow
monitors instrument the program, encoding sensitive values as \emph{faceted values}:
decision trees specifying different views of data according to different possible security labels.
For example, the faceted value \facet{Alice}{\code{true}}{\code{false}}
represents a value that should appear to Alice as \blue{\code{true}} and to
everyone except Alice as \blue{\code{false}}. Faceted execution propagates distinct facets
of a value by extending core linguistic primitives (such as function application).
For example, consider the application \blue{\code{(x true)}} where
\code{x} is \blue{\facet{Alice}{\lambda x.~\code{true}}{\code{not}}}.
Racket's standard function application will fail here because Racket's
\blue{\code{ppapp}} expects a procedure rather than a facet. Instead, the proper way
to interpret function application on faceted values is to distribute the
application over all (in this case, both) facets, producing
\facet{Alice}{\code{true}}{\code{false}}: if Alice is viewing, the application yielded true,
otherwise it yielded false, so both are computed until the value is explicitly observed
with specified permissions. Many other core forms (such
as \code{if}, \code{set!}, etc...) require similar changes to handle faceted
values correctly.

@@

In this paper we present \racets, an implementation of policy-agnostic
programming in Racket via macros. \racets provides facilities for
creating policies and faceting secure data with those
policies. \racets also extends several core forms in Racket to work
with faceted values (our implementation is detailed in
Section~\ref{sec:implementation}). We have used \racets to implement a
small server-based board-game (detailed in Sections~\ref{sec:overview}
and~\ref{sec:implementation}). Relevant related work is presented in
section~\ref{sec:related}. We see \racets as a promising prototype for
policy-agnostic programming in Racket, and conclude with discussion of
future directions in Section~\ref{sec:conclusion}.

@@

The semantics of faceted execution is an extension of the lambda
calculus, leading to a natural question: can we use Racket's macros to
extend Racket to faceted execution? We will see that the answer is
yes, and the translation from the big-step rules is surprisingly
straightforward. This section of our paper describes the design of
\racets, a prototype implementation of faceted execution using Racket
macros. In Section~\ref{sec:conclusion} we remark upon current
directions scaling \racets to the whole of Racket.

>>>>

\emph{Faceted execution} (FE) is a highly expressive 
language semantics enabling policy-agnostic programming. In FE, dynamic information-flow
monitors instrument the program, encoding sensitive values as \emph{faceted values}:
decision trees specifying different views of data according to different possible security labels.
For example, the faceted value \facet{Alice}{\code{true}}{\code{false}}
represents a value that should appear to Alice as \blue{\code{true}} and to
everyone except Alice as \blue{\code{false}}. Faceted execution propagates distinct facets
of a value by extending core linguistic primitives (such as function application).
For example, consider the application \blue{\code{(x true)}} where
\code{x} is \blue{\facet{Alice}{\lambda x.~\code{true}}{\code{not}}}.
Racket's standard function application will fail here because Racket's
\blue{\code{ppapp}} expects a procedure rather than a facet. Instead, the proper way
to interpret function application on faceted values is to distribute the
application over all (in this case, both) facets, producing
\facet{Alice}{\code{true}}{\code{false}}: if Alice is viewing, the application yielded true,
otherwise it yielded false, so both are computed until the value is explicitly observed
with specified permissions. Many other core forms (such
as \code{if}, \code{set!}, etc...) require similar changes to handle faceted
values correctly.

@@

In this paper we present \racets, an implementation of policy-agnostic
programming in Racket via macros~\cite{Kohlbecker:1986, Dybvig:1992}. \racets provides facilities for
creating policies and faceting secure data with those
policies. \racets also extends several core forms in Racket to work
with faceted values (our implementation is detailed in
Section~\ref{sec:implementation}). We have used \racets to implement a
small server-based board-game (detailed in Sections~\ref{sec:overview}
and~\ref{sec:implementation}). Relevant related work is presented in
section~\ref{sec:related}. We see \racets as a promising prototype for
policy-agnostic programming in Racket, and conclude with discussion of
future directions in Section~\ref{sec:conclusion}.

@@

The semantics of faceted execution is an extension of the lambda
calculus, leading to a natural question: can we use Racket's macros~\cite{Kohlbecker:1986, Dybvig:1992} to
extend Racket to faceted execution? We will see that the answer is
yes, and the translation from the big-step rules is surprisingly
straightforward. This section of our paper describes the design of
\racets, a prototype implementation of faceted execution using Racket
macros. In Section~\ref{sec:conclusion} we remark upon current
directions scaling \racets to the whole of Racket.
